---
title: "HW4 Solutions"
author: "Rushal Sohal; github: rsohal"
date: "3/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
```

**1.**

```{r}
weight0 <- ChickWeight %>% filter(Time == 0) %>% group_by(Diet) %>% summarize(num_chicks = n(), avg_weight = mean(weight), sd_weight = sd(weight))
weight0
```


<!-- 5 pts of extra credit: construct a ggplot that shows average weights by diet with an interval (shown as a line) of +- the standard deviation around the averages. -->
```{r}
ggplot(weight0, aes(x=Diet, y=avg_weight)) + geom_point() + geom_errorbar(aes(ymin = avg_weight-sd_weight, ymax = avg_weight+sd_weight)) + ggtitle('Extra Credit: Average Weight by Diet') + xlab('Diet') + ylab('Average Weight')
```



**2.**

```{r}
complete <- ChickWeight %>% group_by(Chick) %>% mutate(num_obs = n()) %>% ungroup() %>% filter(num_obs == 12)
head(complete, n=3)
```


**3.**

```{r}
complete <- complete %>% group_by(Chick) %>% mutate(weightgain = (weight - weight[Time==0]))
head(complete, n=3)
```



**4.**


```{r}
filter(complete, Time==21) %>% ggplot(aes(x = reorder(Diet, weightgain, FUN = median), y = weightgain)) + geom_boxplot() + ylim(0,350) + ggtitle('Weight Gain by Diet for Day 21') + xlab('Diet') + ylab('Weight Gain')
```

Diet 3 has the highest on average weight gain on Day 21 whereas Diet 1 has the least. Diet 4 has the least variation or deviation.



**5.**
```{r}
print(complete %>% filter(Time==21) %>% arrange(desc(weightgain)) %>% head(1))
complete %>% filter(Time==21) %>% arrange(weightgain) %>% head(1)
```

Graphs for weight gain by chick and by diet.
```{r}
complete %>% filter(Time==21) %>% ggplot(aes(x=Chick, y=weightgain)) + geom_point() + ylab('Weightgain') + ggtitle('Weightgain by Chick')
complete %>% filter(Time==21) %>% ggplot(aes(x=Diet, y=weightgain)) + geom_point() + ylab('Weightgain') + ggtitle('Weightgain by Diet')
```

Chick 35 gained the most weight (=332) and was on Diet 3. Chick 24 gained the least (=32) and was on Diet 2.

**6.**

```{r}
complete %>% filter(Time==0) %>% arrange(weight) %>% head(1)
```

```{r}
complete %>% filter(Time==0) %>% arrange(desc(weight)) %>% head(1)
```
```{r}
plot1 <- complete %>% filter(Chick == 27|  Chick == 19) %>% ggplot(aes(x=Time, y=weight, group=Chick, color=Chick)) + geom_line() + geom_point() + ggtitle('Weight over Time')
plot1
```





**7.**

```{r}
tmp <- complete %>% group_by(Time) %>% summarize(avg = mean(weight)) 
plot2 <- complete %>% group_by(Chick) %>% filter(Chick ==27 | Chick == 19) %>% ggplot() + geom_line(aes(x=Time, y=weight, group=Chick, color=Chick))+ geom_line(data=tmp, aes(x=Time, y=avg))
plot2
```

The three trajectories look similar - exponentially increasing. However, the mean trajectory is greater or of higher value compared to highest and lowest weight chicks.



**8.**
Realtion between the median growth trajectory and the chicks with the least and the most weight (at time 0) for all chicks.
```{r}
tmp <- group_by(complete, Time) %>% summarise(median = median(weightgain))
plot2 + geom_smooth(data = tmp, aes(x=Time, y = median, color ="median"), se=F) +ylim(0,300)
```

Overall, median and lowest/highest weight chick follow a similar trend - increasing (exponential)r. On Day 20, all 3 - median, maximum, and minimum - have a similar growth rate.

